Accelerated evolution of convective simulations

### 2018-03-02 tex/bvp_paper.tex
pretty good draft

\begin{abstract}
We present a method for coupling boundary value problems with initial value problems 
in order to achieve Accelerated Evolution (AE) of convective solutions on
dynamical timescales, rather than the long thermal timescale. 
We study this method in the context of \RB convection. 
We demonstrate that the solution reached by AE and Standard Evolution (SE) are
similar, and that this method works at a
large range of supercriticalities.  The AE method is used to achieve converged 
solutions at high supercriticality ($10^7$), and its extensions to more complex 
systems are briefly discussed.
\end{abstract}
\maketitle

@@

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of natural systems.  For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Convective systems with divergent dynamical timescales can now be studied; however,
thermal timescales, which characterize system relaxation, are often much larger than
dynamical timescales, and resolving dynamics in atmospheres which are sufficiently
thermally relaxed remains a challenging problem.
Solar convection is a prime example of this phenomenon.
Dynamical timescales in the solar convective zone are relatively short 
(10 min overturn at solar surface, one month solar
rotation rate) compared to the Sun's Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
evolving the thermal structure of the system in a meaningful fashion using
traditional timestepping techniques alone.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale becomes intractably large
compared to dynamical timescales \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Increasingly turbulent motions 
require finer grid meshes and smaller timesteps
to capture advective dynamics. Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In other systems, 
the approximate state of the evolved solution can be estimated. There, a
set of initial conditions which is close to the evolved state be 
derived analytically \cite{couston&all2017, brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
However, such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for tens to hundreds of buoyancy times \cite{stevens&all2011}, much less the
thousands of buoyancy timescales required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of non-converged
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean thermodynamic profile towards its evolved state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare solutions reached through the accelerated evolution
method to those that have been evolved through a full diffusive timescale, and we
examine select simulations at high Ra which have achieved accelerated evolution. Finally,
in section \ref{sec:extensions}, we discuss extensions of the methods presented here,
and we offer concluding remarks.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that our fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except where it is $\rho = \rho_0(1  - \alpha T_1)$ on the term where
the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, 
acts in the vertical momentum equation.
The equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that the
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
by these choices, one time unit is a freefall time ($L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, where $\angles{}_{x,y}$ represent
a horizontal average, are the basis of the AE method. We measure
both of these quantities early in a simulation, retrieving profiles similar to
those shown in Fig. \ref{fig:time_trace}b.
At early stages in the simulation, these flux profiles are highly asymmetric,
with more flux exiting the atmosphere at the upper boundary than 
the fixed-flux lower boundary is providing.
By calculating the total flux,
F$_{\text{tot}} =$ F$_{\text{conv}} +$ F$_{\text{cond}}$, we derive the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{conv}}(z) = 1$
and $f_{\text{cond}}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\text{cond}}(z) = 1$ and $f_{\text{conv}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

Thus, the AE method is simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the fraction,
as a function of height, by which the flux in the system must be reduced to go from
an initial state (Fig. \ref{fig:time_trace}b) to an evolved state (Fig. \ref{fig:time_trace}d).
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}_{x,y}$, by $\sqrt{\xi}$, such 
that the advected, flux-carrying perturbations in the temperature are appropriately
diminished.  We then solve a boundary value problem including
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}) using F$_{\text{conv, ev}}$
and $\cdot\angles{\bm{u}\times\bm{\omega}}_{x,y}$, after diminishing it by a factor
of $\xi$. We use mixed thermal boundary conditions, as in Eqn. (\ref{eqn:bcs}),
because $F_{\text{bot}}$ is exactly specified, and the temperature profile is pegged
at a reference value for the AE solve.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

@@

%%%%%%%%%%%%
%%%%%%%%%%%
% RESULTS
%%%%%%%%%%%
%%%%%%%%%%%%

\section{Results}
\label{sec:results}
We study evolved solutions achieved through the standard evolution (SE) method 
from convective onset
up to $S = 10^5$ in 2D and $S = 10^4$ in 3D.  These SE runs are compared to
accelerated evolution (AE) runs spanning from onset up to $S = 10^7$ in 2D and $S = 10^4$ in 3D.
For a full list of simulations, we refer the reader to appendix \ref{appendix:run_table}.

We report the time- and volume-averaged values of select measurements of the
evolved solutions in Fig. \ref{fig:parameter_space_comparison}.
The scaling of heat transport in the evolved solution, as quantified by the
Nusselt number, is shown in Fig. \ref{fig:parameter_space_comparison}a.
The volume averaged Nusselt number is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where $\angles{}$ represent a volume average.
In 2D when $S < 10^{3+2/3}$ and in 3D, the evolved system is defined by a clear
value of Nu and the convective heat transport reaches a temporally stationary state.
In 2D and at larger values of $S$, the value of Nu oscillates as a function of time
due to large horizontal oscillations in the convective plume structures.
Our choice of no-slip
boundary conditions prevent the fluid from entering a shearing state
\cite{goluskin&all2014}, but the oscillatory motions which do arise cause the
system to vary between periods of low heat transport and high heat transport.
The SE simulations which span up to $S = 10^5$ exhibit the same horizontally
oscillatory motion as the AE solutions for the same initial conditions. The
scaling of the mean value of Nu is roughly $\text{Nu} \propto \text{Ra}^{1/5}$,
weaker than that reported in similar systems with fixed-T and fixed-flux boundary
conditions \cite{johnston&doering2009}.  We attribute this weaker Nu scaling to
the oscillatory nature of the plumes, which may have been avoided by previous studies
using bootstrapping techniques as initial conditions.

@@

In Fig. \ref{fig:parameter_space_comparison}b, we report the volume-averaged
RMS Reynolds number in the AE solutions, where
$\text{Re} = \angles{|\bm{u}|} / \mathcal{R}$.  This measure scales roughly as
$\text{Re} \propto \text{Ra}^{0.45}$, and shows little variance with time.

In Fig. \ref{fig:parameter_space_comparison}c, we report the volume averaged 
temperature of the AE solutions, with the value at the upper (fixed $T$) boundary removed.
This measurement probes the thickness of the boundary layers of the solutions, and
should show scaling which is inversely proportional to Nu in converged solutions
where fixed-flux boundary conditions are used \cite{otero&all2002}.  We find here
that $(\angles{T} - T_{\text{top}}) \propto \text{Ra}^{-1/5}$, precisely the inverse
scaling of Nu, giving confidence that these solutions are in a converged state.

In Fig. \ref{fig:parameter_space_comparison}d-f, we report the fractional difference
between measurements from the AE solutions and measurements from the SE solutions.
We find that the mean value of Nu from AE solutions is accurate to the values
from SE solutions to within $\sim 1$\%, and the same is true for $\angles{T}$ measurements.
Re measurements show slightly greater error, with AE measurements being on average
$\leq 2$\% away from the SE measurements.

The measurements presented in Fig. \ref{fig:parameter_space_comparison} demonstrate
that the AE method can be powerfully employed in parameter space studies in which
large numbers of simulations are compared in a volume-averaged sense.  We now turn
our examination to a more direct comparison of AE and SE for convection at
$S = 10^5$, as has been introduced in Fig. \ref{fig:time_trace}.

As the AE method primarily serves to adjust the thermodynamic structure of the
solution, we compare the temperature profile attained by AE and SE in 
Fig. \ref{fig:temp_comparison}.  We see that the boundary layer length scale is 
nearly identical between the two solutions (Fig. \ref{fig:temp_comparison}a), but that
the mean temperature in the interior differs by about 0.5\% on average
(Fig. \ref{fig:temp_comparison}c).

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
$T$ is sampled every 0.1 time units for 500 total time units,
and was interpolated onto an evenly spaced grid before sampling.
(c) The percentage difference between the mean temperature profiles as a function of height.
The difference between the mean profiles is very small, O(0.5\%).
(d) The value of the Kolmogorov-Smirnov (KS) statistic, or the difference between
the AE and SE CDFs, as a function of temperature.
The small
difference in the interior temperature 
results in a large difference between the two temperature CDFs near the values
of the temperature modes.  The spread of temperature around the modes, which includes the 
fluctuations that drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

>>>>

\begin{abstract}
We present a method for achieving Accelerated Evolution (AE) of convective simulations.
By probing the dynamics of convection early in simulations and adjusting the
thermal profile of the simulation appropriately, we advance simulations to
evolved states on timescales much shorter than the thermal diffusion timescale.
We study this method in the context of \RB convection. 
The solutions reached by AE similar to those reached by a slow thermal evolution, 
and AE is effective in 2D, 3D, and at low and high supercriticality.  
Extensions to more complex systems are briefly discussed.
\end{abstract}
\maketitle

@@

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of systems in nature.  For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Unfortunately, convective systems generally evolve over the course of thermal
timescales, which are often much larger than
dynamical timescales. Therefore resolving dynamics in atmospheres which are sufficiently
thermally relaxed remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 10$ minutes at the solar surface, 
the Sun rotates roughly once every month)
compared to the Sun's Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale becomes intractably large
compared to dynamical timescales \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In other systems, 
the approximate state of the evolved solution can be estimated. There, a
set of initial conditions which is close to the evolved state can be 
derived analytically \cite{couston&all2017, brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
However, such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for tens to hundreds of freefall timescales \cite{stevens&all2011}, much less the
thousands of freefall times required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of non-converged
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean thermodynamic profile towards its evolved state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from evolution through a full thermal diffusion timescale. Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except where it is $\rho = \rho_0(1  - \alpha T_1)$ on the term where
the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, 
acts in the vertical momentum equation.
The equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig. \ref{fig:time_trace}b.
At early stages in the simulation, these flux profiles are highly asymmetric,
with more flux exiting the atmosphere at the upper boundary than 
the fixed-flux lower boundary is providing.
By calculating the total flux,
F$_{\text{tot}} =$ F$_{\text{conv}} +$ F$_{\text{cond}}$, we derive the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{conv}}(z) = 1$
and $f_{\text{cond}}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\text{cond}}(z) = 1$ and $f_{\text{conv}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

Thus, the AE method is simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the fraction,
as a function of height, by which the flux in the system must be reduced to go from
an initial state (Fig. \ref{fig:time_trace}b) to an evolved state (Fig. \ref{fig:time_trace}d).
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}_{x,y}$, by $\sqrt{\xi}$, such 
that the advected, flux-carrying perturbations in the temperature are appropriately
diminished.  We then solve a 1D boundary value problem including
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}) using F$_{\text{conv, ev}}$
and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$, after diminishing it by a factor
of $\xi$. We use mixed thermal boundary conditions in this study, 
as in Eqn. (\ref{eqn:bcs}),
because $F_{\text{bot}}$ is exactly specified, and the temperature profile is fixed 
at a reference value for the AE solve.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

@@

%%%%%%%%%%%%
%%%%%%%%%%%
% RESULTS
%%%%%%%%%%%
%%%%%%%%%%%%
\vspace{-0.5cm}
\section{Results}
\label{sec:results}
We study evolved SE solutions in $S \in (1, 10^5]$ in 2D and $S \in (1, 10^4]$ in
3D. We compare their properties to
accelerated evolution (AE) runs in $S \in (1, 10^7]$ in 2D and
$S \in (1, 10^4]$ in 3D.
We refer the reader to appendix \ref{appendix:run_table} for a full list of
simulations.

\vspace{5cm}
$\,$\newline
The time- and volume-averaged values of the Nusselt number (Nu), the RMS
Reynolds number (Re), and the mean temperature in AE solutions 
are shown in Fig. \ref{fig:parameter_space_comparison}a-c.
Nu, which quantifies the heat transport in evolved solutions,
is shown as a function of Ra and $S$ in Fig. \ref{fig:parameter_space_comparison}a.
Nu is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $g$, is shown as $\angles{g}$.
When $S < 10^{3+2/3}$ in 2D and for all runs in 3D, 
the evolved system is characterized by a specific
value of Nu, as the convective heat transport reaches a temporally stationary state.
At larger values of $S$ in 2D, the convective plumes consistently oscillate
horizontally, leading to variations in Nu over time as the heat transport is alternately
assisted and diminished by these oscillations.
As a result of our choice of no-slip
boundary conditions, the fluid never enters a full domain shearing state 
\cite{goluskin&all2014}, and the
oscillatory nature of the plumes is stable.
The 2D SE simulations exhibit the same horizontally
oscillatory behavior as the AE solutions for the same initial conditions. The
dynamic nature of the plume structures diminishes the
scaling of the heat transport to $\text{Nu} \propto \text{Ra}^{1/5}$,
which is weaker than classic scaling laws \cite{johnston&doering2009, ahlers&all2009}.
Previous studies in 2D convection may have avoided oscillatory plumes by using
bootstrapping techniques as initial conditions for high $S$ runs.

@@

In Fig. \ref{fig:parameter_space_comparison}b, we report 
$\text{Re} = \angles{|\bm{u}|} / \mathcal{R}$.  This measures the degree of
turbulence in the solution, and scales roughly as
$\text{Re} \propto \text{Ra}^{0.45}$. Each simulation shows little variance in
the value of Re over time.

Fig. \ref{fig:parameter_space_comparison}c shows the scaling of $\angles{T} - T_{\text{top}}$.
The average temperature, $\angles{T}$, is dominated by its value in the isothermal interior,
so this measurements serves as a probe of the temperature jump across the boundary
layers. As a result, we expect this measure to
scaling inversely with Nu in converged solutions
where fixed-flux boundary conditions are used \cite{otero&all2002}.  We find here
that $(\angles{T} - T_{\text{top}}) \propto \text{Ra}^{-1/5}$, precisely the inverse
scaling of Nu.

In Fig. \ref{fig:parameter_space_comparison}d-f, we report the fractional difference
between measurements in the AE and SE solutions.
The mean values of Nu and $\angles{T} - T_{\text{top}}$ 
measured in AE are accurate to SE values to within $\sim 1$\%.
Re measurements show marginally greater error, with AE measurements being 
$\leq 2$\% different from SE measurements.

The measurements presented in Fig. \ref{fig:parameter_space_comparison} demonstrate
that AE can be powerfully employed in parameter space studies in which
large numbers of simulations are compared in a volume-averaged sense.  We now turn
our examination to a more direct comparison of AE and SE for convection at
$S = 10^5$, the time and flux evolution of which was shown in Fig. \ref{fig:time_trace}.
All comparisons that follow for these two runs occur over the times shaded in
green in Fig. \ref{fig:time_trace}a\&c. Measurements are sampled every
0.1 freefall time units for 500 total time units.

As AE is fundamentally a 1D adjustment to the thermodynamic structure of the
solution, we compare the horizontally- and time-averaged temperature profiles 
attained by AE and SE in Fig. \ref{fig:temp_comparison}a.  
The boundary layer width and structure are  
nearly identical between the two solutions, but the
the mean temperature in the isothermal interior differs by roughly 0.5\%
(Fig. \ref{fig:temp_comparison}c).

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
$T$ is sampled every 0.1 time units for 500 total time units,
and was interpolated onto an evenly spaced grid before sampling.
(c) The percentage difference between the mean temperature profiles as a function of height.
The difference between the mean profiles is very small, O(0.5\%).
(d) The value of the Kolmogorov-Smirnov (KS) statistic, or the difference between
the AE and SE CDFs, as a function of temperature.
The small
difference in the interior temperature 
results in a large difference between the two temperature CDFs near the values
of the temperature modes.  The spread of temperature around the modes, which includes the 
fluctuations that drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

The probability distribution functions (PDFs)
of point-by-point temperature measurements are compared for the two runs
in Fig. \ref{fig:temp_comparison}b. To construct these PDFs, 
we interpolate the temperature field
of each of these snapshots onto an evenly spaced grid, determine the
frequency distribution of all measured $T$ values, and then normalize the
distribution such that its integral is unity.  The two PDFs have noticeably
different modes, as is expected from Fig. \ref{fig:temp_comparison}a. 
Over long timescales, the 0.5\% difference between the two profiles would
disappear, as the AE solution evolves to be exactly the SE solution -- this
is evident in the asymmetry of the AE PDF in Fig. \ref{fig:temp_comparison}b and also
the trend of the mean temperature over time in Fig. \ref{fig:temp_comparison}c.
